---
title: "World Demography Dashboard" 
# author: "CID: 02511851"
# date: "5 May, 2024"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    vertical_layout: fill
    theme: yeti
    orientation: rows
runtime: shiny
---

``` {js}
$('.navbar-inverse').removeClass('navbar-inverse').addClass('navbar-default');
```


```{r setup, include=FALSE}
# Load necessary packages. 
library(flexdashboard)
library(highcharter)
library(viridis)
library(tidyverse)
library(countrycode)
library(DT)
library(crosstalk)
library(plotly)
library(dplyr)
library(tibble)
require(shiny)
require(gridExtra)
```

```{r include=FALSE}
population <- read.csv('population-and-demography.csv')
population <-  population[, c("Country.name", 
                              "Year", 
                              "Population", 
                              "Population.of.children.under.the.age.of.5", 
                              "Population.aged.5.to.9.years", 
                              "Population.aged.10.to.14.years", 
                              "Population.aged.15.to.19.years",
                              "population__all__20_24__records",
                              "population__all__25_29__records",
                              "population__all__30_34__records",
                              "population__all__35_39__records", 
                              "population__all__40_44__records",
                              "population__all__45_49__records",
                              "population__all__50_54__records",
                              "population__all__55_59__records",
                              "population__all__60_64__records",
                              "population__all__65_69__records",
                              "population__all__70_74__records",
                              "population__all__75_79__records",
                              "population__all__80_84__records",
                              "population__all__85_89__records",
                              "population__all__90_94__records",
                              "population__all__95_99__records",
                              "Population.older.than.100.years")] %>% 
  rename("Country" = "Country.name",
         "0-4" = "Population.of.children.under.the.age.of.5",
         "5-9" = "Population.aged.5.to.9.years",
         "10-14" = "Population.aged.10.to.14.years",
         "15-19" = "Population.aged.15.to.19.years",
         "20-24" = "population__all__20_24__records",
         "25-29" = "population__all__25_29__records",
         "30-34" = "population__all__30_34__records",
         "35-39" = "population__all__35_39__records",
         "40-44" = "population__all__40_44__records",
         "45-49" = "population__all__45_49__records",
         "50-54" = "population__all__50_54__records",
         "55-59" = "population__all__55_59__records",
         "60-64" = "population__all__60_64__records",
         "65-69" = "population__all__65_69__records",
         "70-74" = "population__all__70_74__records",
         "75-79" = "population__all__75_79__records",
         "80-84" = "population__all__80_84__records",
         "85-89" = "population__all__85_89__records",
         "90-94" = "population__all__90_94__records",
         "95-99" = "population__all__95_99__records",
         "100 + " = "Population.older.than.100.years")

gender_ratio <- read.csv('share-population-female.csv') %>% 
  rename("Country" = "Entity", "Female"="Population..female....of.total.population.")

gender_ratio[,"Male"] <- 100 - gender_ratio[,"Female"]

fertility_rate <- read.csv('children-per-woman-vs-human-development-index.csv')
fertility_rate <- fertility_rate[,c("Entity", "Year", "Fertility.rate...Sex..all...Age..all...Variant..estimates")] %>% 
  rename("Country" = "Entity", 
         "Fertility_rate" = "Fertility.rate...Sex..all...Age..all...Variant..estimates")

data <- inner_join(population, gender_ratio, by = c("Country", "Year")) 
data <- inner_join(data, fertility_rate, by = c("Country", "Year"))

names(data) <- tolower(names(data))

countries <- data[!(data$country %in% c("Low-income countries", 
                                        "Lower-middle-income countries", 
                                        "Upper-middle-income countries", 
                                        "High-income countries", 
                                        "World")), ]

countries[countries$country == "United States", "country"] = "United States of America" 
countries[countries$country == "Serbia", "country"] = "Republic of Serbia"

income <- data[(data$country %in% c("Low-income countries", 
                                    "Lower-middle-income countries", 
                                    "Upper-middle-income countries", 
                                    "High-income countries", 
                                    "World")), ]
```

```{r include=FALSE}
# Create a custom theme for the plots. 
custom_theme <- hc_theme(
  colors = c('#5CACEE', 'green', 'red'),
  chart = list(
         backgroundColor = '#FFFFFF', 
         plotBorderColor = "black"),
  xAxis = list(
         gridLineColor = "C9C9C9", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#C9C9C9", 
         minorGridLineColor = "#C9C9C9", 
         tickColor = "#C9C9C9", 
         title = list(style = list(color = "#333333"))), 
  yAxis = list(
         gridLineColor = "#C9C9C9", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#C9C9C9", 
         minorGridLineColor = "#C9C9C9", 
         tickColor = "#C9C9C9", 
         tickWidth = 1, 
         title = list(style = list(color = "#333333"))),   
  title = list(style = list(color = '#333333', fontFamily = "Lato")),
  subtitle = list(style = list(color = '#666666', fontFamily = "Lato")),
  legend = list(
         itemStyle = list(color = "#333333"), 
         itemHoverStyle = list(color = "#FFF"), 
         itemHiddenStyle = list(color = "#606063")), 
  credits = list(style = list(color = "#666")),
  itemHoverStyle = list(color = 'gray'))
```

<style>
.colored {
  background-color: #eef4fc;
}
.navbar {
  background-color: #5B92E4;
  border-color: #FFFFFF;
}

</style>

<style type="text/css">
.value-box .value {
    color: white;
}
.value-box .caption {
    color: white;
}
</style>


Worldwide
=======================================================================

Row {data-height = 50}
-----------------------------------------------------------------------

### World Population in 2021
```{r}
world_average_fertility_rate <- countries %>%
  select(year, country, population) %>%
  group_by(year) %>%
  summarise(population = sum(population)) 

world_average_fertility_rate <- tail(world_average_fertility_rate$population, 1)

valueBox(world_average_fertility_rate, icon = "fa-globe", color = '#5B92E4')
```

Row {data-height = 600}
-----------------------------------------------------------------------

### World map {.no-title}

```{r}
countries_tibble <- countries %>%
  select(year, country, population) %>%
  group_by(year, country) %>%
  summarise(population = sum(population))

selectInput("worldwide_year", 
            "Year:", 
            choices = unique(countries_tibble$year), 
            selected = unique(countries_tibble$year)[-1])

renderHighchart({
  countries_tibble_data <- reactive({
    subset(countries_tibble, year == input$worldwide_year) 
  })

  
  highchart() %>%
      hc_add_series_map(worldgeojson,
                        countries_tibble_data(),
                        value = "population",
                        joinBy = c("name", "country")) %>% 
      hc_colorAxis(minColor = "#dee9f9", 
                   maxColor = "#243a5b") %>% 
      hc_title(text = paste("Population by Country for Year", 
                            input$worldwide_year)) %>% 
      hc_tooltip(borderWidth = 1.5, headerFormat = "") %>% 
      hc_plotOptions(series = list(stacking = "normal", 
                                   dataLabels = list(enabled = TRUE, 
                                                     format = "{point.y:.0f}"))) %>%
      hc_add_theme(custom_theme)
})
```

Row {.tabset .tabset-fade .colored data-height=350}
-----------------------------------------------------------------------
### Fertility Rate and Population Change

```{r}
world_fertility_rate <- countries %>% 
  select(year, fertility_rate) %>% 
  group_by(year) %>% 
  summarise(fertility_rate = mean(fertility_rate))

output1 <- highchart() %>% 
  hc_add_series(world_fertility_rate, 
                hcaes(x = year, y = fertility_rate, color = fertility_rate), 
                type = "line") %>%  
  hc_tooltip(
    crosshairs = TRUE, 
    borderWidth = 1.5, 
    headerFormat = "", 
    pointFormat = paste("Year: <b>{point.x}</b> Fertility rate: <b>{point.y}</b>")
  ) %>%
  hc_title(text = "Worldwide Fertility Rate by Year") %>% 
  hc_subtitle(text = "1960-2021") %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Fertility rate"), allowDecimals = FALSE) %>%
  hc_legend(enabled = FALSE) %>% 
  hc_add_theme(custom_theme)

world_population <- countries %>%
  select(year, population) %>%
  group_by(year) %>%
  summarise(population = sum(population)) 

output2 <- highchart() %>% 
  hc_add_series(world_population, 
                hcaes(x = year, y = population, color = population), 
                type = "line") %>%
  hc_tooltip(
    crosshairs = TRUE, 
    borderWidth = 1.5, 
    headerFormat = "", 
    pointFormat = paste("Year: <b>{point.x}</b> Population: <b>{point.y}</b>")
  ) %>%
  hc_title(text = "Worldwide Population by Year") %>% 
  hc_subtitle(text = "1960-2021") %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Population"),
           labels = list(formatter = JS("function() { 
                                        return this.value / 1000000000 + 'B'; }")),
           allowDecimals = FALSE
  ) %>%
  hc_legend(enabled = FALSE) %>% 
  hc_add_theme(custom_theme)

splitLayout(cellWidths = c("50%"), output1, output2)
```

### Population Pyramid by Year

```{r}
world_gender_ratio_male <- countries %>% 
  select('year', 'male', '0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + ') %>% 
  group_by(year) %>% 
  summarise('0-4' = round(sum(`0-4` * `male` / 100), 0),
            '5-9' = round(sum(`5-9` * `male` / 100), 0),
            '10-14' = round(sum(`10-14` * `male` / 100), 0),
            '15-19' = round(sum(`15-19` * `male` / 100), 0),
            '20-24' = round(sum(`20-24` * `male` / 100), 0),
            '25-29' = round(sum(`25-29` * `male` / 100), 0),
            '30-34' = round(sum(`30-34` * `male` / 100), 0),
            '35-39' = round(sum(`35-39` * `male` / 100), 0),
            '40-44' = round(sum(`40-44` * `male` / 100), 0),
            '45-49' = round(sum(`45-49` * `male` / 100), 0),
            '50-54' = round(sum(`50-54` * `male` / 100), 0),
            '55-59' = round(sum(`55-59` * `male` / 100), 0),
            '60-64' = round(sum(`60-64` * `male` / 100), 0),
            '65-69' = round(sum(`65-69` * `male` / 100), 0),
            '70-74' = round(sum(`70-74` * `male` / 100), 0),
            '75-79' = round(sum(`75-79` * `male` / 100), 0),
            '80-84' = round(sum(`80-84` * `male` / 100), 0),
            '85-89' = round(sum(`85-89` * `male` / 100), 0),
            '90-94' = round(sum(`90-94` * `male` / 100), 0),
            '95-99' = round(sum(`95-99` * `male` / 100), 0),
            '100 + ' = round(sum(`100 + ` * `male` / 100), 0)) %>% 
  pivot_longer(!year, names_to = "Age", values_to = "Male")

world_gender_ratio_female <- countries %>% 
  select('year', 'female', '0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + ') %>% 
  group_by(year) %>% 
  summarise('0-4' = round(sum(`0-4` * `female` / 100), 0) ,
            '5-9' = round(sum(`5-9` * `female` / 100), 0),
            '10-14' = round(sum(`10-14` * `female` / 100), 0),
            '15-19' = round(sum(`15-19` * `female` / 100), 0),
            '20-24' = round(sum(`20-24` * `female` / 100), 0),
            '25-29' = round(sum(`25-29` * `female` / 100), 0),
            '30-34' = round(sum(`30-34` * `female` / 100), 0),
            '35-39' = round(sum(`35-39` * `female` / 100), 0),
            '40-44' = round(sum(`40-44` * `female` / 100), 0),
            '45-49' = round(sum(`45-49` * `female` / 100), 0),
            '50-54' = round(sum(`50-54` * `female` / 100), 0),
            '55-59' = round(sum(`55-59` * `female` / 100), 0),
            '60-64' = round(sum(`60-64` * `female` / 100), 0),
            '65-69' = round(sum(`65-69` * `female` / 100), 0),
            '70-74' = round(sum(`70-74` * `female` / 100), 0),
            '75-79' = round(sum(`75-79` * `female` / 100), 0),
            '80-84' = round(sum(`80-84` * `female` / 100), 0),
            '85-89' = round(sum(`85-89` * `female` / 100), 0),
            '90-94' = round(sum(`90-94` * `female` / 100), 0),
            '95-99' = round(sum(`95-99` * `female` / 100), 0),
            '100 + ' = round(sum(`100 + ` * `female` / 100)), 0) %>% 
  pivot_longer(!year, names_to = "Age", values_to = "Female")

world_gender_ratio <- inner_join(world_gender_ratio_male, 
                                 world_gender_ratio_female, 
                                 by = c("year", "Age")) 

categories <- c(
  '0-4', '5-9', '10-14', '15-19',
  '20-24', '25-29', '30-34', '35-39', '40-44',
  '45-49', '50-54', '55-59', '60-64', '65-69',
  '70-74', '75-79', '80-84', '85-89', '90-94',
  '95-99', '100 + '
)  

renderHighchart({
  pyramid_data_male <- reactive({
    subset(world_gender_ratio_male, year == input$worldwide_year) %>% 
      select(-year) %>% 
      mutate(Age = factor(Age, levels = rev(levels(Age))))
  })
  pyramid_data_female <- reactive({
    subset(world_gender_ratio_female, year == input$worldwide_year) %>% 
      select(-year) %>% 
      mutate(Age = factor(Age, levels = rev(levels(Age))))
  })
  
  formatLargeNumber <- JS("function() {
      var absValue = Math.abs(this.value);
      if (absValue >= 1000000) {
          return (absValue / 1000000).toFixed(1) + 'M';
      } else if (absValue >= 1000) {
          return (absValue / 1000).toFixed(1) + 'K';
      } else {
          return absValue;
      }
  }")
  
  highchart() %>% 
    hc_chart(type = "bar") %>% 
    hc_title(text = paste("Population Pyramid for Year", input$worldwide_year)) %>% 
    hc_xAxis(categories = categories, label = list()) %>% 
    hc_yAxis(
      title = list(text = "Population"),
      labels = list(
        formatter = formatLargeNumber
      )
    ) %>% 
    hc_add_series(name = "Male", 
                  data = pyramid_data_male(), 
                  type = "bar", 
                  hcaes(x = Age, y = -Male)) %>% 
    hc_add_series(name = "Female", 
                  data = pyramid_data_female(), 
                  type = "bar", 
                  hcaes(x = Age, y = Female)) %>% 
    hc_tooltip(
      formatter = JS("function() {
                      return '<b>' + this.series.name + '</b><br/>' +
                             'Age: <b>' + this.point.category + '</b><br/>' +
                             'Population: <b>' +
                             Highcharts.numberFormat(Math.abs(this.point.y), 0) + '</b>';
                  }")
    ) %>%
    hc_plotOptions(series = list(stacking = "normal", 
                                 dataLabels = list(enabled = FALSE, 
                                                   format = "{point.y:.0f}"))) %>%
    hc_colors(c("#5B92E4", "#E55B92")) %>%  
    hc_add_theme(custom_theme) 
})

```

By Income
=======================================================================

Row {data-height=650}
-----------------------------------------------------------------------
### Worldwide population pyramid by income {.no-title}

```{r}
income_gender_ratio_male <- income %>% 
  select('year', 'country', 'male', '0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + ') %>% 
  group_by(year, country) %>% 
  summarise('0-4' = sum(`0-4` * `male` / 100) ,
            '5-9' = sum(`5-9` * `male` / 100),
            '10-14' = sum(`10-14` * `male` / 100),
            '15-19' = sum(`15-19` * `male` / 100),
            '20-24' = sum(`20-24` * `male` / 100),
            '25-29' = sum(`25-29` * `male` / 100),
            '30-34' = sum(`30-34` * `male` / 100),
            '35-39' = sum(`35-39` * `male` / 100),
            '40-44' = sum(`40-44` * `male` / 100),
            '45-49' = sum(`45-49` * `male` / 100),
            '50-54' = sum(`50-54` * `male` / 100),
            '55-59' = sum(`55-59` * `male` / 100),
            '60-64' = sum(`60-64` * `male` / 100),
            '65-69' = sum(`65-69` * `male` / 100),
            '70-74' = sum(`70-74` * `male` / 100),
            '75-79' = sum(`75-79` * `male` / 100),
            '80-84' = sum(`80-84` * `male` / 100),
            '85-89' = sum(`85-89` * `male` / 100),
            '90-94' = sum(`90-94` * `male` / 100),
            '95-99' = sum(`95-99` * `male` / 100),
            '100 + ' = sum(`100 + ` * `male` / 100)) %>% 
  pivot_longer(c('0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + '), names_to = "Age", values_to = "Male")

income_gender_ratio_female <- income %>% 
  select('year', 'country', 'female', '0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + ') %>% 
  group_by(year, country) %>% 
  summarise('0-4' = sum(`0-4` * `female` / 100) ,
            '5-9' = sum(`5-9` * `female` / 100),
            '10-14' = sum(`10-14` * `female` / 100),
            '15-19' = sum(`15-19` * `female` / 100),
            '20-24' = sum(`20-24` * `female` / 100),
            '25-29' = sum(`25-29` * `female` / 100),
            '30-34' = sum(`30-34` * `female` / 100),
            '35-39' = sum(`35-39` * `female` / 100),
            '40-44' = sum(`40-44` * `female` / 100),
            '45-49' = sum(`45-49` * `female` / 100),
            '50-54' = sum(`50-54` * `female` / 100),
            '55-59' = sum(`55-59` * `female` / 100),
            '60-64' = sum(`60-64` * `female` / 100),
            '65-69' = sum(`65-69` * `female` / 100),
            '70-74' = sum(`70-74` * `female` / 100),
            '75-79' = sum(`75-79` * `female` / 100),
            '80-84' = sum(`80-84` * `female` / 100),
            '85-89' = sum(`85-89` * `female` / 100),
            '90-94' = sum(`90-94` * `female` / 100),
            '95-99' = sum(`95-99` * `female` / 100),
            '100 + ' = sum(`100 + ` * `female` / 100)) %>% 
  pivot_longer(c('0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + '), names_to = "Age", values_to = "Female")

income_gender_ratio <- inner_join(income_gender_ratio_male, 
                                  income_gender_ratio_female, 
                                  by = c("year", "country", "Age")) 

fluidRow(
  column(width = 6,
         selectInput("income_year", 
                     "Year:", 
                     choices = unique(income_gender_ratio$year), 
                     selected = unique(income_gender_ratio$year)[-1])
  ),
  column(width = 6,
         selectInput("income_level", 
                     "Income Level:", 
                     choices = unique(income_gender_ratio$country), 
                     selected = unique(income_gender_ratio$country)[1])
  )
)

categories <- c(
  '0-4', '5-9', '10-14', '15-19',
  '20-24', '25-29', '30-34', '35-39', '40-44',
  '45-49', '50-54', '55-59', '60-64', '65-69',
  '70-74', '75-79', '80-84', '85-89', '90-94',
  '95-99', '100 + '
)  

renderHighchart({
  income_gender_ratio_male_reactive <- reactive({
    subset(income_gender_ratio_male, 
           year == input$income_year & country == input$income_level) %>% 
      mutate(Age = factor(Age, levels = rev(levels(Age)))) # Reorder age groups
  })
  
  income_gender_ratio_female_reactive <- reactive({
    subset(income_gender_ratio_female, 
           year == input$income_year & country == input$income_level) %>% 
      mutate(Age = factor(Age, levels = rev(levels(Age)))) # Reorder age groups
  })
  
  formatLargeNumber <- JS("function() {
      var absValue = Math.abs(this.value);
      if (absValue >= 1000000) {
          return (absValue / 1000000).toFixed(1) + 'M';
      } else if (absValue >= 1000) {
          return (absValue / 1000).toFixed(1) + 'K';
      } else {
          return absValue;
      }
  }")

  highchart() %>% 
    hc_chart(type = "bar") %>% 
    hc_title(text = paste("Population Pyramid in", input$income_level, "for Year", input$income_year)) %>% 
    hc_xAxis(categories = categories, label = list()) %>% 
    hc_yAxis(title = list(text = "Population"),
             labels = list(formatter = formatLargeNumber)) %>% 
    hc_add_series(name = "Male", 
                  data = income_gender_ratio_male_reactive(), 
                  type = "bar", 
                  hcaes(x = Age, y = -Male, name = "Male")) %>% 
    hc_add_series(name = "Female",
                  data = income_gender_ratio_female_reactive(), 
                  type = "bar", 
                  hcaes(x = Age, y = Female, name = "Female")) %>% 
    hc_tooltip(
      formatter = JS("function() {
                      return '<b>' + this.series.name + '</b><br/>' +
                             'Age: <b>' + this.point.category + '</b><br/>' +
                             'Population: <b>' +
                             Highcharts.numberFormat(Math.abs(this.point.y), 0) + '</b>';
                  }")
    ) %>%
    hc_plotOptions(series = list(stacking = "normal", 
                                 dataLabels = list(enabled = FALSE, 
                                                   format = "{point.y:.0f}"))) %>%
    hc_colors(c("#5B92E4", "#E55B92")) %>%  
    hc_add_theme(custom_theme)
})
```

Column {data-height = 400 .colored}
-----------------------------------------------------------------------
### Worldwide population by income {.no-title}

```{r}
# Create tibble for our line plot (World population by year).  
fertility_rate_income <- income %>% 
  select(year, country, fertility_rate) 

income_output1 <- highchart() %>% 
  hc_add_series(fertility_rate_income, 
                hcaes(x = year, 
                      y = fertility_rate, 
                      group = country, 
                      color = country), 
                type = "line") %>% 
  
  hc_tooltip(crosshairs = TRUE, 
             borderWidth = 1.5, 
             headerFormat = "", 
             pointFormat = paste("Year: <b>{point.x}</b> <br> 
                                 Type: <b>{point.country}</b> <br> 
                                 Fertility rate: <b>{point.y}</b>")) %>%
  hc_title(text = "Worldwide fertility rate by year") %>% 
  hc_subtitle(text = "1960-2021") %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Fertility rate"), allowDecimals = FALSE) %>%
  hc_add_theme(custom_theme)  

world_population_income <- income %>%
  select(year, country, population)

income_output2 <- highchart() %>% 
  hc_add_series(world_population_income, 
                hcaes(x = year, 
                      y = population, 
                      group = country, 
                      color = country), 
                type = "line") %>%
  hc_tooltip(crosshairs = TRUE, 
             borderWidth = 1.5, 
             headerFormat = "", 
             pointFormat = paste("Year: <b>{point.x}</b> <br> 
                                 Type: <b>{point.country}</b> <br> 
                                 Population: <b>{point.y}</b>")) %>%
  hc_title(text = "Worldwide population by year") %>% 
  hc_subtitle(text = "1960-2021") %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Population"),
           labels = list(
             formatter = JS("function() { return this.value / 1000000000 + 'B'; }")
             ),
           allowDecimals = FALSE
  ) %>%
  hc_add_theme(custom_theme)

splitLayout(cellWidths = c("50%"), income_output1, income_output2)

```

By Country
=======================================================================

Row
-----------------------------------------------------------------------

### Population pyramid {.no-title}

```{r}
world_gender_ratio_male_2 <- countries %>%
  select('year', 'country', 'male', '0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + ') %>%
  group_by(year, country) %>%
  summarise('0-4' = sum(`0-4` * `male` / 100),
            '5-9' = sum(`5-9` * `male` / 100),
            '10-14' = sum(`10-14` * `male` / 100),
            '15-19' = sum(`15-19` * `male` / 100),
            '20-24' = sum(`20-24` * `male` / 100),
            '25-29' = sum(`25-29` * `male` / 100),
            '30-34' = sum(`30-34` * `male` / 100),
            '35-39' = sum(`35-39` * `male` / 100),
            '40-44' = sum(`40-44` * `male` / 100),
            '45-49' = sum(`45-49` * `male` / 100),
            '50-54' = sum(`50-54` * `male` / 100),
            '55-59' = sum(`55-59` * `male` / 100),
            '60-64' = sum(`60-64` * `male` / 100),
            '65-69' = sum(`65-69` * `male` / 100),
            '70-74' = sum(`70-74` * `male` / 100),
            '75-79' = sum(`75-79` * `male` / 100),
            '80-84' = sum(`80-84` * `male` / 100),
            '85-89' = sum(`85-89` * `male` / 100),
            '90-94' = sum(`90-94` * `male` / 100),
            '95-99' = sum(`95-99` * `male` / 100),
            '100 + ' = sum(`100 + ` * `male` / 100)) %>% 
  pivot_longer(c('0-4', '5-9', '10-14', '15-19',
                 '20-24', '25-29', '30-34', '35-39', '40-44',
                 '45-49', '50-54', '55-59', '60-64', '65-69',
                 '70-74', '75-79', '80-84', '85-89', '90-94',
                 '95-99', '100 + '), 
               names_to = "Age", 
               values_to = "Male")

world_gender_ratio_female_2 <- countries %>%
  select('year', 'country', 'female', '0-4', '5-9', '10-14', '15-19',
         '20-24', '25-29', '30-34', '35-39', '40-44',
         '45-49', '50-54', '55-59', '60-64', '65-69',
         '70-74', '75-79', '80-84', '85-89', '90-94',
         '95-99', '100 + ') %>%
  group_by(year, country) %>%
  summarise('0-4' = sum(`0-4` * `female` / 100) ,
            '5-9' = sum(`5-9` * `female` / 100),
            '10-14' = sum(`10-14` * `female` / 100),
            '15-19' = sum(`15-19` * `female` / 100),
            '20-24' = sum(`20-24` * `female` / 100),
            '25-29' = sum(`25-29` * `female` / 100),
            '30-34' = sum(`30-34` * `female` / 100),
            '35-39' = sum(`35-39` * `female` / 100),
            '40-44' = sum(`40-44` * `female` / 100),
            '45-49' = sum(`45-49` * `female` / 100),
            '50-54' = sum(`50-54` * `female` / 100),
            '55-59' = sum(`55-59` * `female` / 100),
            '60-64' = sum(`60-64` * `female` / 100),
            '65-69' = sum(`65-69` * `female` / 100),
            '70-74' = sum(`70-74` * `female` / 100),
            '75-79' = sum(`75-79` * `female` / 100),
            '80-84' = sum(`80-84` * `female` / 100),
            '85-89' = sum(`85-89` * `female` / 100),
            '90-94' = sum(`90-94` * `female` / 100),
            '95-99' = sum(`95-99` * `female` / 100),
            '100 + ' = sum(`100 + ` * `female` / 100)) %>% 
  pivot_longer(c('0-4', '5-9', '10-14', '15-19',
                 '20-24', '25-29', '30-34', '35-39', '40-44',
                 '45-49', '50-54', '55-59', '60-64', '65-69',
                 '70-74', '75-79', '80-84', '85-89', '90-94',
                 '95-99', '100 + '), 
               names_to = "Age", 
               values_to = "Female")

world_gender_ratio_2 <- inner_join(world_gender_ratio_male_2, 
                                   world_gender_ratio_female_2, 
                                   by = c("year", "country", "Age")) 

categories_2 <- c('0-4', '5-9', '10-14', '15-19',
                  '20-24', '25-29', '30-34', '35-39', '40-44',
                  '45-49', '50-54', '55-59', '60-64', '65-69',
                  '70-74', '75-79', '80-84', '85-89', '90-94',
                  '95-99', '100 + ')

fluidRow(
  column(width = 6,
         selectInput("by_country_year", 
                     "Year:", 
                     choices = unique(world_gender_ratio_2$year))
  ),
  column(width = 6,
         selectInput("by_country_country", 
                     "Country:", 
                     choices = unique(world_gender_ratio_2$country))
  )
)

renderHighchart({
  pyramid_data_male_2 <- reactive({
    subset(world_gender_ratio_male_2, 
           year == input$by_country_year & country == input$by_country_country) %>% 
      mutate(Age = factor(Age, levels = rev(levels(Age))))
  })
  pyramid_data_female_2 <- reactive({
    subset(world_gender_ratio_female_2, 
           year == input$by_country_year & country == input$by_country_country) %>%
      mutate(Age = factor(Age, levels = rev(levels(Age))))
  })

  formatLargeNumber <- JS("function() {
      var absValue = Math.abs(this.value);
      if (absValue >= 1000000) {
          return (absValue / 1000000).toFixed(1) + 'M';
      } else if (absValue >= 1000) {
          return (absValue / 1000).toFixed(1) + 'K';
      } else {
          return absValue;
      }
  }")
  
  highchart() %>% 
    hc_chart(type = "bar") %>% 
    hc_title(text = paste("Population Pyramid in", 
                          input$by_country_country, 
                          "for Year", 
                          input$by_country_year)) %>% 
    hc_xAxis(categories = categories_2, label = list()) %>% 
    hc_yAxis(title = list(text = "Population"),
             labels = list(formatter = formatLargeNumber)) %>% 
    hc_add_series(name = "Male", 
                  data = pyramid_data_male_2(), 
                  type = "bar", 
                  hcaes(x = Age, y = -Male)) %>% 
    hc_add_series(name = "Female", 
                  data = pyramid_data_female_2(), 
                  type = "bar", 
                  hcaes(x = Age, y = Female)) %>% 
    hc_plotOptions(series = list(stacking = "normal")) %>%
    hc_tooltip(formatter = JS("function() {
                                return '<b>' + this.series.name + '</b><br/>' +
                                'Age: <b>' + this.point.category + '</b><br/>' +
                                'Population: <b>' + 
                                Highcharts.numberFormat(Math.abs(this.point.y), 0) + '</b>';
                              }")) %>%
    hc_colors(c("#5B92E4", "#E55B92")) %>%  
    hc_add_theme(custom_theme)
})
```


```{r}
selectInput("Year3", 
            "Year:", 
            choices = unique(world_gender_ratio_2$year))

selectInput("Country", 
            "Country:", 
            choices = unique(world_gender_ratio_2$country))
```

Column
-----------------------------------------------------------------------

### Fertility Rate by Country {.no-title}
```{r}
fertility_rate_by_country <- countries %>%
  select(year, country, fertility_rate) 

renderHighchart({
  fertility_rate_by_country_reactive <- reactive({
    subset(fertility_rate_by_country, 
           country == input$by_country_country)
  })
  
  
highchart() %>% 
  hc_add_series(fertility_rate_by_country_reactive(), 
                hcaes(x = year, 
                      y = fertility_rate, 
                      color = fertility_rate), 
                type = "line") %>%
  
  hc_tooltip(crosshairs = TRUE, 
             borderWidth = 1.5, 
             headerFormat = "", 
             pointFormat = paste("Year: <b>{point.x}</b> <br> 
                                 Fertility rate: <b>{point.y}</b>")) %>%
  hc_title(text = paste("Fertility rate in", 
                        input$by_country_country, 
                        "by year")) %>% 
  hc_subtitle(text = "1960-2021") %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Population"),
           allowDecimals = FALSE) %>%
  hc_legend(enabled = FALSE) %>% 
  hc_add_theme(custom_theme)
})

```

### Population Change by Country {.no-title}

```{r}
population_change_by_country <- countries %>%
  select(year, country, population) 

renderHighchart({
  population_change_by_country_reactive <- reactive({
    subset(
      population_change_by_country, 
      country == input$by_country_country
    )
  })
  
highchart() %>% 
  hc_add_series(
    population_change_by_country_reactive(), 
    hcaes(x = year, 
    y = population, 
    color = population), 
    type = "line"
  ) %>%
  hc_tooltip(
    crosshairs = TRUE, 
    borderWidth = 1.5, 
    headerFormat = "", 
    pointFormat = paste("Year: <b>{point.x}</b> <br> Population: <b>{point.y}</b>")
  ) %>%
  hc_title(
    text = paste(
      "Population Change in", 
      input$by_country_country, 
      "by year"
    )
  ) %>% 
  hc_subtitle(text = "1960-2021") %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Population"),
           allowDecimals = FALSE) %>%
  hc_legend(enabled = FALSE) %>% 
  hc_add_theme(custom_theme)
})
```


Row
-----------------------------------------------------------------------

### Population in the selected year

```{r}
selected_year_population <- countries %>%
  select(year, country, population)

renderValueBox({
  selected_year_population <- subset(
    selected_year_population, 
    country == input$by_country_country & year == input$by_country_year
  ) %>% 
    select(population) %>% 
    as.numeric
  
  valueBox(selected_year_population, icon = "fa-solid fa-users")
})

```

### Fertility rate in the selected year

```{r}
selected_year_fertility_rate <- countries %>%
  select(year, country, fertility_rate)

renderValueBox({
  selected_year_fertility_rate <- 
    subset(
      selected_year_fertility_rate, 
      country == input$by_country_country & year == input$by_country_year
    ) %>% 
    select(fertility_rate) %>% 
    as.numeric
  
  valueBox(selected_year_fertility_rate, icon = "fas fa-baby")
})
```

### Gender ratio in the selected year

```{r}
selected_year_gender_ratio <- countries %>% 
  select(year, country, male, female)

renderValueBox({
  selected_year_gender_ratio <- 
    subset(
      selected_year_gender_ratio, 
      country == input$by_country_country & year == input$by_country_year
    ) %>% 
    select(male, female) 
  selected_year_gender_ratio_male <- selected_year_gender_ratio$male %>% round(2)
  selected_year_gender_ratio_female <- selected_year_gender_ratio$female %>% round(2)

  selected_year_gender_ratio_total <- selected_year_gender_ratio_male + selected_year_gender_ratio_female
  
  selected_year_gender_ratio_male <- selected_year_gender_ratio_male + (100 - selected_year_gender_ratio_total) * (selected_year_gender_ratio_male / selected_year_gender_ratio_total)
  
  selected_year_gender_ratio_female <- selected_year_gender_ratio_female + (100 - selected_year_gender_ratio_total) * (selected_year_gender_ratio_female / selected_year_gender_ratio_total)
  
  selected_year_gender_ratio_male <- as.character(selected_year_gender_ratio_male)
  selected_year_gender_ratio_female <- as.character(selected_year_gender_ratio_female)

  selected_year_gender_ratio <- paste(
    selected_year_gender_ratio_male, 
    selected_year_gender_ratio_female, 
    sep = ":"
  )
  
  valueBox(selected_year_gender_ratio, icon = "fa-solid fa-venus-mars")
})
```



About {data-icon="fa-info-circle"}
=======================================================================

### {}

**References:**  

The datasets used to create the dashboard can be found at:

* https://ourworldindata.org/population-growth
* https://ourworldindata.org/age-structure
* https://ourworldindata.org/gender-ratio
* https://ourworldindata.org/fertility-rate

and were compiled from data from the following sources: 

* United Nations, World Population Prospects (2022), United Nations Development Program, Human Development Report (2024).

* United Nations Development Program, Human Development Report (2024). 
